---
title: "Robinson_B1"
author: "Daniel"
date: "2023-11-02"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, include=FALSE}
library(devtools)
library(tidyverse)
library(testthat)
library(palmerpenguins)
```

###Excercise 1: Building a Function

In my thesis data, I have several categories of repeatedly sampled field observations. For example, I recorded canopy density at the same 50 locations every week over the course of 4 months. For one of my thesis chapters, I assess spatial differences in forest structure, and am not really interested in the week to week change in data, as much as the site to site differences. For this chapter I need to average the data for each location across each observation period (i.e. I average all 12 observations of canopy cover at fixed point = A from Jan to Apr to get a seasonal average of canopy cover for point A). I will perform this operation on several different variables at various sample sizes and with varying observation amounts, but with the same grouping structure. **In this project, I will make a function consisting of the dplyr commands necessary to group and summarize repeated measure data at different sample sizes and different observation scales.**

``` {r defining function: avg_obs()}
#' Average repeated observations: this function groups data = data by a categorical group variable = group_var to determine the average of a given numeric variable = measure_var
#'
#' @param data a dataframe or tibble (I designited this param as "data" as that seems to be industry standard)
#' @param group_var a categorical variable (I designated this param "group_var" as opposed to var1 or something less descriptive as group_var helps to clarify the variables type as categorical and its role in the function)
#' @param measure_var a numeric variable (I designated this param "measure_var" as opposed to var2 or something less descriptive as measure_var helps to clarify the variables type as numeric and its role in the function)
#'
#' @return a dataframe or tibble of measured variable averages grouped by group variable
#' @export
#'
#' @examples
avg_obs <- function(data, group_var, measure_var, na.rm = TRUE, ...) { #added na.rm = TRUE to ignore NA as the default
  result <- data %>%
    group_by({{ group_var }}) %>%  # Group the data by the specified variable
    summarize(mean_value = mean({{ measure_var }}, na.rm = na.rm, ...))  # Calculate the mean of the specified measure variable
  
  return(result)
}

```

We will run a few examples with simple mock dataframes to check the function by eye.

```{r mock testing function: avg_obs()}
#Creating sample dataframe 
data <- data.frame(
  id = c(1, 1, 2, 2, 3, 3), #Sample categorical variable for group_var
  value = c(10, 15, 20, 25, 30, 35) #Sample numeric variable for measure_var
)

result <- avg_obs(data, group_var = id, measure_var = value)
result

```

```{r mock testing the funciton with NAs: avg_obs()}
#Creating sample dataframe with NAs to assess robustness
data <- data.frame(
  id = c(1, 1, 2, 2, 3, 3, 3, NA, NA), #Sample categorical variable for group_var
  value = c(10, 15, 20, 25, 30, 35, NA, NA, 40) #Sample numeric variable for measure_var
)

result <- avg_obs(data, group_var = id, measure_var = value)
result
```

And now we will test our function on the penguins dataset.

``` {r mock testing the function with penguins dataset}
#Checking out the penguin data 
head(penguins)

#Island and species are good candidates for out grouping variable and bill_depth_mm or any of the numeric variables can be averaged:
result <- avg_obs(penguins, species, bill_length_mm)
```


The function looks like its working okay with the mock tests, but now we will write formal tests for the function using expect_() family of functions and test_that() from the testthat package.

```{r formally testing the function with expect_()}
test_that("data input and output classes are correct", {
  # Sample dataframe from above
  penguins

  # Call the function from above
  result <- avg_obs(penguins, species, bill_length_mm)

  # Use expect_ family of functions to define expectations for variable class
  expect_true(is.numeric(result$mean_value))
  expect_true(is.factor(penguins$species))
  expect_true(is.numeric(penguins$bill_length_mm))
})
  
```

Now we will run a test to make sure out mean values are making sense given out original data.

```{r testthat}
#Use expect_equal to test that the number of rows matches number of species and test that the mean values of the result make sense given the values from the original data 
test_that("output makes sense numerically", {
  expect_equal(nrow(result), n_distinct(penguins$species))  # Expect the number of rows in the result to match the number of species
  for (species_group in unique(penguins$species)) {   # Creating loop to check that mean_value of result is less than the maximum value for each group
    max_value_group <- max(penguins$bill_length_mm[penguins$species == species_group], na.rm = TRUE)
    expect_true(all(result$mean_value[result$species == species_group] < max_value_group))
  }
})

```

Lastly, we will run an easy test just to makes sure there are no NAs in mean_value data

```{r testthat}
test_that("No NAs in mean_value", {
  expect_false(anyNA(result$mean_value))
})
```
